What We Eval in the Shadows

### 2021-08-09 paper/main.tex
Explain ... pattern

Assertion packages such as \code{assertthat} provide functions to check pre and
post conditions and produce friendly error messages on failures. These
conditions are written as expressions which are evaluated using \eval. Consider
the \code{see_\if} function shown below. It accepts the conditions as arguments
to the \code{...} parameter. The code of these conditions is extracted using
\code{eval(substitute(alist(...)))}. The conditions are then evaluated
sequentially by the expression \code{eval(assertion, env)}, inside a
\code{tryCatch} block. If a condition evaluates to \code{FALSE}, the function
returns immediately with an error message, leaving the remaining conditions
unevaluated.

@@

\subsection{Miscellaneous}
A common use of \eval is the extraction of unevaluated text of arguments passed
to \code{...} (vararg). For a non-vararg parameter \code{x}, the expressions
\code{substitute(x)} can be used to extract the argument text. However, for \code{...}
arguments, the expression \code{eval(quote(substitute(list(...))))} or its variants
are used. The \eval call evaluates the quoted expression
\code{substitute(list(...))}. This returns a \code{list} of unevaluated argument
expressions. The \code{hasGroupNumber} function from the \code{lattice} package is one
of the many users of this pattern.

>>>>

An unexpected use of \eval is the extraction of unevaluated text of arguments
passed to a \code{...} parameter. Consider the definition of the \code{NVL}
function from the \code{statnet.common} package. This function returns the first
argument that is non \code{NULL}.

\begin{lstlisting}
NVL <- function(...) {
    for(e in eval(substitute(alist(...))))
        x <- eval(e, parent.frame())
        if(!is.null(x)) break
    }
    x
}
\end{lstlisting}
The expression \code{eval(substitute(alist(...)))} is used to extract the
unevaluated text of the arguments passed to \code{NVL}.
\code{substitute(alist(...))} returns an expression representing a call to the
\code{alist} function with \code{...} replaced by the arguments supplied to
\code{NVL}. Evaluating this expression returns a \code{list} whose arguments are
the unevaluated argument expressions which are evaluated sequentially for their
result.

Assertion packages such as \code{assertthat} provide functions to check pre and
post conditions and produce friendly error messages on failures. These
conditions are written as expressions which are evaluated using \eval. Consider
the \code{see\_if} function shown below. It accepts the conditions as arguments
to the \code{...} parameter. The code of these conditions is extracted using
\code{eval(substitute(alist(...)))}. The conditions are then evaluated
sequentially by the expression \code{eval(assertion, env)}, inside a
\code{tryCatch} block. If a condition evaluates to \code{FALSE}, the function
returns immediately with an error message, leaving the remaining conditions
unevaluated.

@@

\subsection{Miscellaneous}
